---
title: "COMPASS_TEMPESTX_SoilFluxes_8100"
author: "Zoe Read"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
---
## Goals of this code: 
1. Read in 8100 files - with chamber ID, start time, end time, V1-V4
2. Separate 8100 file into each flux 
3. Extract metadata from each flux file into a metadata sheet

##Setup: 
```{r Setup}
library(dplyr)
library(stringr)
library(purrr)
library(tibble)
library(readr)
library(R.utils)
```

## Pull zipped 81x files into temporary data folder
```{r, include=FALSE}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
zip_contents <- unzip("8100_data/Test_FW_Autochamber_Data.zip", list = TRUE)

# Keep only 8100 files from freshwater
metadat_files <- zip_contents$Name[
  grepl("81x$", zip_contents$Name, ignore.case = TRUE)]

# Extract the 8100 files
unzip(
  "8100_data/Test_FW_Autochamber_Data.zip",
  files = metadat_files,
  exdir = tmp_dir
)

# List extracted 8100 files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "81x$",
  full.names = TRUE,
  recursive = TRUE
)

file.list

```

##Create separate files in a  temporary directory: 
```{r}

##Split file into multiple files, one for each chamber run and export to temporary directory 
tmp_dir # view temporary directory

##Need to create a function that will read in all the autochamber files and separate them! 

for (i in file.list) { 
  lines <- readLines(i)

# create file counter
file_counter <- 1
current_chunk <- c()

# go through every line and create a new chunk every time there is "LI-8100"
for (line in lines) {
  # If the line contains "8100", save the current chunk and start a new file
  if (grepl("LI-8100", line)) {
    # Write to file if it is not empty
    if (length(current_chunk) > 0) {
      output_name <- paste0(tmp_dir, "/8100_split_file_", file_counter, ".txt")
      writeLines(current_chunk, output_name)
      file_counter <- file_counter + 1
      current_chunk <- c() # Reset chunk
    }
  }
  
  # append the line to the current chunk
  current_chunk <- c(current_chunk, line)
}

# write the very last chunk of lines to a file
if (length(current_chunk) > 0) {
  output_name <- paste0(tmp_dir, "8100_split_file_", file_counter, ".txt")
  writeLines(current_chunk, output_name)
}
}
```

##Code from fluxfinder functions: 
```{r}

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "txt$",
  full.names = TRUE,
  recursive = TRUE
)

ffi_read_LI78x0 <- function(file, model) {

  dat_raw <- readLines(file)

  # Make sure this is the desired model
  file_model <- substr(dat_raw[1], 1, 7)
  if(file_model != model) stop("This does not appear to be a ", model, " file!")

  # Save the observation number
  obs <- trimws(gsub("Obs#:\t", "", dat_raw[7], fixed = TRUE))
    # Save the observation length
  obs_length <- trimws(gsub("Observation Length:\t", "", dat_raw[10], fixed = TRUE))
  # Save the machine serial number
  sn <- trimws(gsub("Serial Number:\t", "", dat_raw[4], fixed = TRUE))
  # Parse the collar label from the header
  label <- trimws(gsub("Label:\t", "", dat_raw[9], fixed = TRUE))
  # Parse the collar number from the header
  chamber <- trimws(gsub("Chamber:\t", "", dat_raw[14], fixed = TRUE))
  # Parse the chamber volume from the header
  Vtotal <- trimws(gsub("Vtotal:\t", "", dat_raw[25], fixed = TRUE))
  # Parse the dead band from the footer
  DeadBand <- trimws(gsub("Dead Band:\t", "", dat_raw[length(dat_raw)-2], fixed = TRUE))
  # Parse the closing time from the header and use it to make a collar field
  TimeClosing <- trimws(gsub("TimeClosing:\t", "", dat_raw[length(dat_raw)-1], fixed = TRUE))
  
  # These files have 30 header lines, the column names in line 31,
  # and then the column units in line 7. We only want the names and then data
  total_lines = countLines(file)
  lines_to_skip <- 30
  lines_to_read <- total_lines - 30 - 25
  
  dat_read <- read.delim(
    file = file,
    skip = lines_to_skip,
    nrows = lines_to_read,
    sep = "\t",
    header = TRUE)

  # Get first and last timestamp
  start_time <- first(dat_read$Date)
  end_time <- last(dat_read$Date)
  
  # Get average of environmental variables (V1-V4)
  Tcham <- mean(dat_read$Tcham, na.rm = TRUE)
  Pressure <- mean(dat_read$Pressure, na.rm = TRUE)
  V1_avg <- mean(dat_read$V1, na.rm = TRUE)
  V2_avg <- mean(dat_read$V2, na.rm = TRUE)
  V3_avg <- mean(dat_read$V3, na.rm = TRUE)
  V4_avg <- mean(dat_read$V4, na.rm = TRUE)
  
  tmp <- data.frame(
  SN = sn,
  label = label,
  chamber = chamber, 
  Vtotal = Vtotal, 
  obs_length = obs_length, 
  DeadBand = DeadBand, 
  TimeClosing = TimeClosing, 
  start_time = start_time, 
  end_time = end_time,
  Tcham = Tcham, 
  Pressure = Pressure, 
  V1_avg = V1_avg,
  V2_avg = V2_avg,
  V3_avg = V3_avg,
  V4_avg = V4_avg
)

  # ffi_message(basename(file), ": read ", nrow(dat_read), " rows of ", label, " data, ",
  #             start_time, " to ", end_time)

  return(tmp)
}

ffi_read_LI8100 <- function(file) {
  ffi_read_LI78x0(file, "LI-8100")
}

alldat1 <- lapply(file.list, ffi_read_LI8100)

# Combine into one data.table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)


```

##Match the 8100 metadata to the collar metadata
```{r}

collar_metadata <- read.csv("design/collar_map.csv")
collar_metadata$Collar <- as.character(collar_metadata$Collar)

alldat_collars <- alldat %>% 
   left_join(collar_metadata, by=c("label"="Collar"))
 
```


##Write out parsed 8100 data: 
```{r}

folder_name <- "Processed_Data"

if (!dir.exists(folder_name)) {
  dir.create(folder_name)
}

#export as a csv to the folder that you want 
write.csv(alldat_collars, file="Processed_Data/COMPASS_TEMPESTX_SoilFluxes_8100_Metadata_processed.csv", row.names = FALSE)           #Change file name

```

##Remove temporary directory
```{r}
##Remove temporary directory
unlink(tmp_dir, recursive = TRUE)
```

