---
title: "COMPASS_Synoptic_SoilFluxes_2025_8100"
author: "Zoe Read"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
---
## Goals of this code: 
1. Read in 8100 metadata file - with chamber ID, start time, V1-V4, and soil temperature
2. Separate 8100 file into each flux and append the metadata to the flux information 
3. Match with 7810 and 7820 data

Need to add code to parse multiple 8100 files at once
Add code to select the first and last timestamp (probably just with min and max?) and average V1-V4. 

##Setup: 
```{r Setup}
library(dplyr)
library(stringr)
library(purrr)
library(tibble)
library(readr)
library(R.utils)
```

## Pull zipped 81x files into temporary data folder
```{r, include=FALSE}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
zip_contents <- unzip("8100_data/Test_FW_Autochamber_Data.zip", list = TRUE)

# Keep only JSON files
metadat_files <- zip_contents$Name[
  grepl("81x$", zip_contents$Name, ignore.case = TRUE)&
  grepl("FRESH", zip_contents$Name, ignore.case = TRUE)]

# Extract only the JSON files
unzip(
  "8100_data/Test_FW_Autochamber_Data.zip",
  files = metadat_files,
  exdir = tmp_dir
)

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "FRESH.*\\.81x$",
  full.names = TRUE,
  recursive = TRUE
)

head(file.list)

```

##Create separate files in a  temporary directory: 
```{r}

##Split file into multiple files, one for each chamber run and export to temporary directory 
# Create a temporary directory
# tmp_dir <- tempdir()
tmp_dir # view temporary directory

##Need to create a function that will read in all the autochamber files and separate them! 

# # Read file
# file <- "C:\\Users\\readz\\AppData\\Local\\Temp\\RtmpG66gtm/Test_FW_Autochamber_Data/FW_Test_20260520_8100_Files/FRESHWATER/TEMPEST_FRESH-202611410.81x" 

for (i in file.list) { 
  lines <- readLines(i)

# create file counter
file_counter <- 1
current_chunk <- c()

# go through every line and create a new chunk every time there is "LI-8100"
for (line in lines) {
  # If the line contains "8100", save the current chunk and start a new file
  if (grepl("LI-8100", line)) {
    # Write to file if it is not empty
    if (length(current_chunk) > 0) {
      output_name <- paste0(tmp_dir, "/8100_split_file_", file_counter, ".txt")
      writeLines(current_chunk, output_name)
      file_counter <- file_counter + 1
      current_chunk <- c() # Reset chunk
    }
  }
  
  # append the line to the current chunk
  current_chunk <- c(current_chunk, line)
}

# write the very last chunk of lines to a file
if (length(current_chunk) > 0) {
  output_name <- paste0(tmp_dir, "8100_split_file_", file_counter, ".txt")
  writeLines(current_chunk, output_name)
}
}
```

##Code from fluxfinder functions: 
```{r}

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "txt$",
  full.names = TRUE,
  recursive = TRUE
)


ffi_read_LI78x0 <- function(file, model) {

  dat_raw <- readLines(file)

  # Make sure this is the desired model
  file_model <- substr(dat_raw[1], 1, 7)
  if(file_model != model) stop("This does not appear to be a ", model, " file!")

  # Save the observation number
  obs <- trimws(gsub("Obs#:\t", "", dat_raw[7], fixed = TRUE))
    # Save the observation length
  obs_length <- trimws(gsub("Observation Length:\t", "", dat_raw[10], fixed = TRUE))
  # Save the machine serial number
  sn <- trimws(gsub("Serial Number:\t", "", dat_raw[4], fixed = TRUE))
  # Parse the collar label from the header
  label <- trimws(gsub("Label:\t", "", dat_raw[9], fixed = TRUE))
  # Parse the collar number from the header
  chamber <- trimws(gsub("Chamber:\t", "", dat_raw[14], fixed = TRUE))
  # Parse the chamber volume from the header
  Vtotal <- trimws(gsub("Vtotal:\t", "", dat_raw[25], fixed = TRUE))
  # Parse the dead band from the footer
  DeadBand <- trimws(gsub("Dead Band:\t", "", dat_raw[length(dat_raw)-2], fixed = TRUE))
  # Parse the closing time from the header and use it to make a collar field
  TimeClosing <- trimws(gsub("TimeClosing:\t", "", dat_raw[length(dat_raw)-1], fixed = TRUE))
  
  # These files have 30 header lines, the column names in line 31,
  # and then the column units in line 7. We only want the names and then data
  total_lines = countLines(file)
  lines_to_skip <- 30
  lines_to_read <- total_lines - 30 - 25
  
  dat_read <- read.delim(
    file = file,
    skip = lines_to_skip,
    nrows = lines_to_read,
    sep = "\t",
    header = TRUE)

  # Get first and last timestamp
  start_time <- first(dat_read$Date)
  end_time <- last(dat_read$Date)
  
  # Get average of environmental variables (V1-V4)
  V1_avg <- mean(dat_read$V1, na.rm = TRUE)
  V2_avg <- mean(dat_read$V2, na.rm = TRUE)
  V3_avg <- mean(dat_read$V3, na.rm = TRUE)
  V4_avg <- mean(dat_read$V4, na.rm = TRUE)
  
  tmp <- data.frame(
  SN = sn,
  label = label,
  chamber = chamber, 
  Vtotal = Vtotal, 
  obs_length = obs_length, 
  DeadBand = DeadBand, 
  TimeClosing = TimeClosing, 
  start_time = start_time, 
  end_time = end_time,
  V1_avg = V1_avg,
  V2_avg = V2_avg,
  V3_avg = V3_avg,
  V4_avg = V4_avg
)

  # ffi_message(basename(file), ": read ", nrow(dat_read), " rows of ", label, " data, ",
  #             start_time, " to ", end_time)

  return(tmp)
}

ffi_read_LI8100 <- function(file) {
  ffi_read_LI78x0(file, "LI-8100")
}

alldat1 <- lapply(file.list, ffi_read_LI8100)

# Combine into one data.table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)


```

##Write out parsed 8100 data: 
```{r}

folder_name <- "8100_Processed_Data"

if (!dir.exists(folder_name)) {
  dir.create(folder_name)
}

#export as a csv to the folder that you want 
write.csv(alldat, file="8100_Processed_Data/8100_processed_data_test1.csv", row.names = FALSE)           #Change file name

##Remove temporary directory
unlink(tmp_dir, recursive = TRUE)
```

