---
title: "COMPASS_TEMPEST_Autochamber_SoilFlux_CO2_20260604"
author: "Zoe Read & S. Wilson"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
---
## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
library(patchwork)
library(hms)
```

## Pull in Raw Datafiles
```{r, echo = FALSE}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
#zip_contents <- unzip("7810_data/concentrations/Test_FW_Autochamber_Data.zip", list = TRUE)
zip_contents <- unzip(here::here("8100_data", "Autochamber_DataDownload_20260604.zip"), list = TRUE)

# Keep only 8100 files from freshwater
CO2_files <- zip_contents$Name[
  grepl("data$", zip_contents$Name, ignore.case = TRUE)& #8100 files
  grepl("TG10", zip_contents$Name, ignore.case = TRUE)] #Freshwater autochamber test

# CO2_files

# Extract the 8100 files
unzip(
  here::here("8100_data", "Autochamber_DataDownload_20260604.zip"),
  #"7810_data/concentrations/Test_FW_Autochamber_Data.zip",
  files = CO2_files,
  exdir = tmp_dir
)

# List extracted 8100 files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "TG10.*\\.data$",
  full.names = TRUE,
  recursive = TRUE
)

head(file.list)

#Read in all the datafiles 

#Control 7810: TG10-01712
#Saltwater 7810: TG10-01746
#Freshwater 7810: TG10-01725
ctrl_file <- file.list[grepl("TG10-01712", file.list)]
alldat_ctrl <- ffi_read_LI7810(ctrl_file)

fw_file <- file.list[grepl("TG10-01725", file.list)]
alldat_fw <- ffi_read_LI7810(fw_file)

sw_file <- file.list[grepl("TG10-01746", file.list)]
alldat_sw <- ffi_read_LI7810(sw_file)


#combine all the files into one 
#alldat1 <- lapply(file.list, ffi_read_LI7810)
#alldat <- data.table::rbindlist(alldat1, fill=TRUE)

#Read in just the first file: 
#alldat <- ffi_read_LI7810(file.list[2]) #freshwater 

```

## Read in  metadata file and separate by plot
```{r, echo = FALSE, message=FALSE}

#read in metadatfile
#md <- "Processed_Data/COMPASS_TEMPESTX_SoilFluxes_8100_Metadata_processed.csv"
md <- (here::here("Processed_Data", "COMPASS_TEMPESTX_SoilFluxes_8100_Metadata_processed_20260404.csv"))

#metadat <- read_excel(md)
metadat <- read.csv(md) 


#Control metadata 
ctrl_metadat <- metadat %>%
  filter(Plot == "Control") %>%
  mutate(
    # Ensure it's a date-time object
    start_time = ymd_hms(start_time), 
    # Extract date
    Date = as.character(date(start_time)),  
      #Date = as.Date(Date , format = "%Y-%m-%d"),
    # Extract time
    Time = format(start_time, "%H:%M:%S"),
    #Parses the string as "minutes:seconds" and extracts the total duration in seconds from the period
    total_seconds = as.numeric(ms(obs_length)), 
    #Add unique plot ID for each timestamp 
    unique_ID = paste("Collar", label, start_time, sep = "_"), 
    Vtotal_m3 = Vtotal/(100^3)) 

#make some changes to the start times??? 
ctrl_metadat <- ctrl_metadat %>%
  mutate(Time = as_hms(Time)) %>%
  mutate(
    Time = if_else(label %in% c(15, 16),
      Time - 10,
      Time) ) %>%
  mutate(Time = as.character(Time)) %>%
  mutate(total_seconds = as.numeric(total_seconds)) %>%
  mutate(total_seconds = if_else(label %in% c(15, 16),
      total_seconds - 30,
      total_seconds) )


#freshwater metadata 
fw_metadat <- metadat %>%
  filter(Plot == "Fresh") %>%
  mutate(
    # Ensure it's a date-time object
    start_time = ymd_hms(start_time), 
    # Extract date
    Date = as.character(date(start_time)),         
    # Extract time
    Time = format(start_time, "%H:%M:%S"), 
    #Parses the string as "minutes:seconds" and extracts the total duration in seconds from the period
    total_seconds = as.numeric(ms(obs_length)), 
    #Add unique plot ID for each timestamp 
    unique_ID = paste("Collar", label, start_time, sep = "_"), 
    Vtotal_m3 = Vtotal/(100^3)) 

#saltwater metadata 
sw_metadat <- metadat %>%
  filter(Plot == "Salt") %>%
  mutate(
    # Ensure it's a date-time object
    start_time = ymd_hms(start_time), 
    # Extract date
    Date = as.character(date(start_time)),         
    # Extract time
    Time = format(start_time, "%H:%M:%S"), 
    #Parses the string as "minutes:seconds" and extracts the total duration in seconds from the period
    total_seconds = as.numeric(ms(obs_length)), 
    #Add unique plot ID for each timestamp 
    unique_ID = paste("Collar", label, start_time, sep = "_"), 
    Vtotal_m3 = Vtotal/(100^3)) 


```

######## Control MATCHING ##############

## Match Control metadata to raw data
```{r, echo = FALSE}

#match the metadata to the concentration data 
alldat_ctrl$metadat_row <- ffi_metadata_match(
  data_timestamps = alldat_ctrl$TIMESTAMP,
  start_dates = ctrl_metadat$Date,
  start_times = ctrl_metadat$Time,
  obs_lengths = ctrl_metadat$total_seconds) 
# head(alldat)

#Set plot, temp, volume, etc. here 
alldat_ctrl$Plot <- ctrl_metadat$unique_ID[alldat_ctrl$metadat_row]
alldat_ctrl$Volume <- ctrl_metadat$Vtotal_m3[alldat_ctrl$metadat_row]
alldat_ctrl$final_temp <- ctrl_metadat$Tcham[alldat_ctrl$metadat_row] #from 8100 file
alldat_ctrl$V2_avg <- ctrl_metadat$V2_avg[alldat_ctrl$metadat_row] 
alldat_ctrl$V3_avg <- ctrl_metadat$V3_avg[alldat_ctrl$metadat_row] 
alldat_ctrl$V4_avg <- ctrl_metadat$V4_avg[alldat_ctrl$metadat_row] 

ctrl_metadat$metadat_row <- seq_len(nrow(ctrl_metadat))


### Need to deal with weird temperatures - probably take the one from the closest reasonable measurement 

#remove the rows that don't match to metadata 
alldat_ctrl <- alldat_ctrl %>%
filter(!is.na(metadat_row))

#can remove any row that have NA's in the CO2 column 
alldat_ctrl <- alldat_ctrl %>%
filter(!is.na(CO2))

```


######## FRESHWATER MATCHING ##############

## Match Fresh metadata to raw data
```{r, echo = FALSE}

fresh_alldat <- alldat_fw

#match the metadata to the concentration data 
fresh_alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = fresh_alldat$TIMESTAMP,
  start_dates = fw_metadat$Date,
  start_times = fw_metadat$Time,
  obs_lengths = fw_metadat$total_seconds) 
# head(alldat)

#Set plot, temp, volume, etc. here 
fresh_alldat$Plot <- fw_metadat$unique_ID[fresh_alldat$metadat_row]
fresh_alldat$Volume <- fw_metadat$Vtotal_m3[fresh_alldat$metadat_row]
fresh_alldat$final_temp <- fw_metadat$Tcham[fresh_alldat$metadat_row] #from 8100 file
fresh_alldat$V2_avg <- fw_metadat$V2_avg[fresh_alldat$metadat_row] 
fresh_alldat$V3_avg <- fw_metadat$V3_avg[fresh_alldat$metadat_row] 
fresh_alldat$V4_avg <- fw_metadat$V4_avg[fresh_alldat$metadat_row] 

fw_metadat$metadat_row <- seq_len(nrow(fw_metadat))


### Need to deal with weird temperatures - probably take the one from the closest reasonable measurement 

#remove the rows that don't match to metadata 
fresh_alldat <- fresh_alldat %>%
filter(!is.na(metadat_row))

#can remove any row that have NA's in the CO2 column 
fresh_alldat <- fresh_alldat %>%
filter(!is.na(CO2))


```


############### Saltwater Plot ############################

## Match Salt metadata to raw data
```{r, echo = FALSE}

#
salt_alldat <- alldat

#match the metadata to the concentration data 
salt_alldat$metadat_row <- ffi_metadata_match(
  data_timestamps = salt_alldat$TIMESTAMP,
  start_dates = salt_metadat$Date,
  start_times = salt_metadat$Time,
  obs_lengths = salt_metadat$total_seconds) 
# head(alldat)

#Set plot, temp, volume, etc. here 
salt_alldat$Plot <- salt_metadat$unique_ID[salt_alldat$metadat_row]
salt_alldat$Volume <- salt_metadat$Vtotal_m3[salt_alldat$metadat_row]
salt_alldat$final_temp <- salt_metadat$Tcham[salt_alldat$metadat_row] #from 8100 file
salt_alldat$V2_avg <- salt_metadat$V2_avg[salt_alldat$metadat_row] 
salt_alldat$V3_avg <- salt_metadat$V3_avg[salt_alldat$metadat_row] 
salt_alldat$V4_avg <- salt_metadat$V4_avg[salt_alldat$metadat_row] 

salt_metadat$metadat_row <- seq_len(nrow(salt_metadat))


### Need to deal with weird temperatures - probably take the one from the closest reasonable measurement 

#remove the rows that don't match to metadata 
salt_alldat <- salt_alldat %>%
filter(!is.na(metadat_row))

#can remove any row that have NA's in the CO2 column 
salt_alldat <- salt_alldat %>%
filter(!is.na(CO2))

```


## Combine all the matched data back into one dataframe 
```{r}

alldat <-  alldat_ctrl #bind_rows(ctrl_alldat, fresh_alldat, salt_alldat)



```


## Calculate fluxes 
```{r, echo = FALSE, message = FALSE, warning=FALSE}
#change the units
alldat$CO2_umol <- ffi_ppm_to_umol(alldat$CO2, 
                                   volume = alldat$Volume # m3
                                   )    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CO2_umol_m2 <- alldat$CO2_umol / 0.0318

fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CO2_umol_m2", 
                             dead_band = 25) #, c0_band = 5?
# head(fluxes)

#take a look at the fluxes we calculated with R2 
ggplot(fluxes, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CO2 flux (nmol/m2/s)")+ 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) 

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(fluxes, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(fluxes, lin_r.squared < .90, select = Plot:TIMESTAMP_max))

#negative CO2 fluxes
fluxes_negative <- as.data.frame(subset(fluxes,  lin_flux.estimate < 0, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(fluxes$lin_flux.estimate, 0.05) 
high <- quantile(fluxes$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(fluxes, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(fluxes, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)


```

## Check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r, echo=FALSE}

tobechecked <- c()

colnames <- c("Plot")

#Loop to run plots above 
for(P in tobechecked){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  print(p1)
}

#list of the ones that I changed and checked and looked good? 
fixedplots <- c()
  

```


## Spot check a few fluxes with R2 > 0.9
```{r, echo=FALSE, message=FALSE, warning=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Empty list to store plots
plot_list <- list()

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P)
  
  p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  
  plot_list[[length(plot_list) + 1]] <- p1
}

### number of plots per page
plots_per_page <- 2

# split plots into pages
plot_pages <- split(
  plot_list,
  ceiling(seq_along(plot_list) / plots_per_page)
)

# print each page
for(i in seq_along(plot_pages)) {
  
  print(
    wrap_plots(plot_pages[[i]], ncol = 2)
  )
  
}

####################
```

## Take a look at the bad fluxes <0.9 
```{r, echo=FALSE, message=FALSE, warning=FALSE}

# If a flux was started too early/late, change start time or Obs_length in metadata file and note n the note column
plot_list <- list()

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

#Loop to run five random plots subset above 
for(P in bad_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  
  plot_list[[length(plot_list) + 1]] <- p1
}

### number of plots per page
plots_per_page <- 2

# split plots into pages
plot_pages <- split(
  plot_list,
  ceiling(seq_along(plot_list) / plots_per_page)
)

# print each page
for(i in seq_along(plot_pages)) {
  
  print(
    wrap_plots(plot_pages[[i]], ncol = 2)
  )
  
}

####################

#Notes on the bad ones 

##Fluxes that have low R aquared but are okay


##Fluxes to remove: 



##Fluxes to change: 


##Already add notes to metadata for the ones I fixed

##Add qaqc flags 

#Fluxes that need to be removed
# remove_flags <- c("MSM_TR_4_202509", "SWH_TR_1_202509", "SWH_TR_2_202509", "MSM_TR_5_202507")
# fluxes$qaqc_flag <- NA
# fluxes <- fluxes %>%
#     mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% remove_flags, "messy flux, remove", qaqc_flag))
# 
# #Any other fluxes that had low R squared but are okay otherwise
# fluxes <- fluxes %>%
#     mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% bad_plots, "low r squared, but checked and flux is okay", qaqc_flag))

###########################


```

## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE, message=FALSE, warning=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

#Loop to run five random plots subset above 
for(P in tail_plots){ 
  r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  
  plot_list[[length(plot_list) + 1]] <- p1
}

### number of plots per page
plots_per_page <- 2

# split plots into pages
plot_pages <- split(
  plot_list,
  ceiling(seq_along(plot_list) / plots_per_page)
)

# print each page
for(i in seq_along(plot_pages)) {
  
  print(
    wrap_plots(plot_pages[[i]], ncol = 2)
  )
  
}

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Look at the negative CO2 fluxes
```{r, echo=FALSE, message=FALSE, warning=FALSE}

### pull out negative fluxes and put them in a list 
neg_plots <- unique(fluxes_negative$Plot)

#Loop to run five random plots subset above 
for(P in neg_plots){ 
 r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  
  plot_list[[length(plot_list) + 1]] <- p1
}

### number of plots per page
plots_per_page <- 2

# split plots into pages
plot_pages <- split(
  plot_list,
  ceiling(seq_along(plot_list) / plots_per_page)
)

# print each page
for(i in seq_along(plot_pages)) {
  
  print(
    wrap_plots(plot_pages[[i]], ncol = 2)
  )
  
}

####################Negative CO2 fluxes
# add a flag for the fluxes that are negative for CO2 because they should be positive
# fluxes <- fluxes %>%
#     mutate(qaqc_flag = ifelse(is.na(qaqc_flag) & Plot %in% neg_plots, "negative CO2 flux, remove", qaqc_flag))

```

## Plot all fluxes
```{r, echo=FALSE, message=FALSE, warning=FALSE}

#ffi_qaqc(fluxes, Plot, output_file = "qaqc.html", output_dir = getwd(), open_output = TRUE)

### pull a random sample of good fluxes to check and plot 
all_plots <- unique(fluxes$Plot)

#Loop to run five random plots subset above 
for(P in all_plots){ 
 r1 <- subset(alldat, Plot == P)
  p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth(color="red")
  
  plot_list[[length(plot_list) + 1]] <- p1
}

### number of plots per page
plots_per_page <- 2

# split plots into pages
plot_pages <- split(
  plot_list,
  ceiling(seq_along(plot_list) / plots_per_page)
)

# print each page
for(i in seq_along(plot_pages)) {
  
  print(
    wrap_plots(plot_pages[[i]], ncol = 2)
  )
  
}

####################


```


## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(fluxes, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(fluxes, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(fluxes, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Pull together dataframe and export for use elsewhere 
```{r, echo=FALSE}

#head(fluxes1)

####Already flagged poor fluxes, negative fluxes, and low r squared fluxes that are actually okay

#make new columns with the plot; zone; replicate 
IDs <- data.frame(do.call('rbind', strsplit(as.character(fluxes$Plot),'_',fixed=TRUE)))
colnames(IDs) <- c("NA",  "label", "StartTime")
# head(IDs)

#rejoin them to the dataframe
allflux <- cbind(IDs, fluxes)

##Add in notes from metadata - add this in when we fix the fluxes
# metadat$Notes <- NA
# metadat_notes <- metadat[, c("label", "Notes")]
# allflux_notes <- allflux %>% 
#   left_join(metadat_notes, by=c("label"="label"))

#export as a csv to the folder that you want 
# setwd("S:/Biogeochemistry/COMPASS Synoptics/Samples & Data/Fluxes/Macrophyte GHG Fluxes/Final Data")        #Change wd 

allflux$`NA`<- NULL
allflux$`NA`<- NULL

write.csv(allflux, file="Processed_Data/COMPASS_TEMPESTX_SoilFluxes_7810_CO2_processed.csv")           #Change file name

```

